---
title: "PA1_template"
author: "Vikash"
date: "September 20, 2015"
output: html_document
---
Reproducible Research: Peer Assessment 1

Loading and preprocessing the data

```{r loading data}
getwd()
setwd("C:/Users/Vikash/Desktop/Assignments/RProgramming")
file1 <- read.csv("activity.csv", header = TRUE, stringsAsFactors = FALSE )
str(file1)
```
Date present in Strcture of Data is not appropriate as it in Character, hence changing format of date variable

```{r date}
file1$date <- as.Date(file1$date)
str(file1)
```
Counting number of steps for missing values

```{r counting missing steps}
sum(is.na(file1$steps))
```

What is mean and median total number of steps taken per day with histogram and bar plot?
```{r removing invalid values }
file1_valid <- file1[which(!is.na(file1$steps)),]
perday <- tapply(file1_valid$steps, file1_valid$date, sum)
hist(perday, 10,  main = "Total number of steps taken per day", xlab = "")
steps_date <- aggregate(steps ~ date, file1_valid, sum)
head(steps_date)
barplot(steps_date$steps, names.arg = steps_date$date, ylim = c(0, 25000), xlab = "date", ylab = "sum(steps)",)
mean(perday)
median(perday)
mean(steps_date$steps)
median(steps_date$steps)
```

What is the average daily activity pattern?
```{r daily activity pattern}
steps_interval <- aggregate(steps ~ interval, file1_valid, mean)
plot(steps_interval, type = 'l',xlab = "5-Minute-Interval", 
    main = "Daily Activity Pattern", ylab = "Average number of steps")
```

The 5-minute interval contains the maximum number of steps

```{r maximum}
steps_interval$interval[which.max(steps_interval$steps)]
```

Imputing missing values

```{r imputting}
activity.clean <- merge(file1, steps_date, by="date", suffixes=c("", ".mean"))
nas <- is.na(activity.clean$steps)
activity.clean$steps[nas] <- activity.clean$steps.mean[nas]
activity.clean <- activity.clean[, c(1:3)]
head(activity.clean)
```

Histogram of the total number of steps taken each day

```{r total steps after inputting}
steps_date <- aggregate(steps ~ date, activity.clean, sum)
barplot(steps_date$steps, names.arg = steps_date$date, ylim=c(0, 25000), 
        xlab="date", ylab="sum(steps)" , )
```

Mean and median of total steps taken
```{r mean/median total steps}
mean(steps_date$steps)
median(steps_date$steps)
```

Are there differences in activity patterns between weekdays and weekends?
Add new factor variable dayType with 2 levels – “weekday” and “weekend”
```{r factor weekdays/weekends}
dayType <- function(dates) {
  f <- function(date) {
    if (weekdays(date) %in% c("Saturday", "Sunday")) {
      "weekend"
    }
    else {
      "weekday"
    }
  }
  sapply(dates, f)
}

file1$dayType <- as.factor(dayType(file1$date))
str(file1)
```

```{r panel plot}
steps_interval <- aggregate(steps ~ interval + dayType, file1, mean)
library(lattice)
xyplot(steps ~ interval | dayType, data=steps_interval, layout=c(2,1), type='l')
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
